Declarative runtime reconfiguration of pulverised distributed systems \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-09 paper-2023-acsos-pulverization.tex
related survey on reconfiguration verification
docs: description of the first dsl

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\author{\IEEEauthorblockN{1\textsuperscript{st} Danilo Pianini}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
danilo.pianini@unibo.it}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Roberto Casadei}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
roby.casadei@unibo.it}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Nicolas Farabegoli}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Mirko Viroli}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
email address or ORCID}
}

@@

@@

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{listing-style}

\usepackage{cleveref}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\author{\IEEEauthorblockN{1\textsuperscript{st} Danilo Pianini}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
danilo.pianini@unibo.it}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Roberto Casadei}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
roby.casadei@unibo.it}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Nicolas Farabegoli}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
nicolas.farabegoli@unibo.it}
\and
\IEEEauthorblockN{4\textsuperscript{th} Mirko Viroli}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
email address or ORCID}
}

@@

The framework allows the system configuration via two distinct DSLs, one for defining the system's devices and their components,
and one for defining the runtime configuration of the system.

The first domain-specific language allows to define the system in terms of
\emph{devices}, \emph{components} and which \emph{capabilities} requires to be executed.
%
In this way, infrastructural and deployment aspects are abstracted, allowing the focus to be on functional aspects of the system.
%
The DSL enables the definition of the type of devices that took part in the system,
and for each device can be specified which components constitute it and where those component must run on (e.g. which host must execute the component).
%
The execution of a component is constrained to hosts with predetermined characteristics:
the framework models the constraint with the \texttt{Capability} concept,
namely the property that a host must to ``implement'' to run the component.
%
Different capabilities can be specified for each component to provide a more flexible and scalable deployment of the system.

The~\Cref{lst:system-configuration} provides an example of the usage of the domain-specific language where is defined a system composed of two kind of devices.
%
The DSL returns the system's configuration that will be used by the runtime.

One key-point of the domain-specific language can be observed in the definition of the \texttt{device-2} device:
the \emph{Behaviour} component can be deployed on hosts that implement the \texttt{HighCPU} and \texttt{EmbeddedDevice} capabilities,
in this way the behaviour can be moved (at runtime) from a host to another one that implements the same capabilities.
%
The association between a component and a single capability do not necessarily imply that the component can be executed only on a single host,
but it is the way to specify the minimum requirements for the host that will execute the component.

\lstinputlisting[
    float=ht,
    language=Kotlin,
    caption={TODO},
    label=lst:system-configuration,
]{listings/SystemConfig.kt}

\lstinputlisting[
    float=ht,
    language=Kotlin,
    caption={TODO},
    label=lst:runtime-configuration,
]{listings/RuntimeConfig.kt}

@@

A recent survey on 
 formal techniques for verifying the correctness of reconfigurations
 in component-based distributed software systems is provided in~\cite{coullon2023swreconfig}.
